### 2025-07-08 mixed-models.qmd
20250701 - tweaks
20250708 - track chunk execution times
20250708 - packages and citations

```{r}
library("lme4")
library("lmerTest")
library("effectsize")
library("MuMIn")
library("emmeans")
library("sjstats")
library("mgcv")
library("AICcmodavg")
library("bbmle")
library("rstan")
library("brms")
library("cmdstanr") # todo: install.packages("cmdstanr", repos = c("https://stan-dev.r-universe.dev", getOption("repos"))); cmdstanr::check_cmdstan_toolchain(); cmdstanr::install_cmdstan()
library("fitdistrplus")
library("performance")
library("parallel")
library("parallelly")
library("broom.mixed")
library("tidybayes")
library("plotly")
library("viridis")
library("tidyverse")
```

@@

>>>>

```{r}
library("lme4")
library("lmerTest")
library("effectsize")
library("MuMIn")
library("emmeans")
library("sjstats")
library("mgcv")
library("AICcmodavg")
library("bbmle")
library("rstan")
library("brms")
library("cmdstanr") # todo: install.packages("cmdstanr", repos = c("https://stan-dev.r-universe.dev", getOption("repos"))); cmdstanr::check_cmdstan_toolchain(); cmdstanr::install_cmdstan()
library("fitdistrplus")
library("performance")
library("parallelly")
library("broom.mixed")
library("tidybayes")
library("plotly")
library("viridis")
library("tidyverse")
```

@@

#### Multicollinearity {#sec-mixedModelsMulticollinearity}

As with [multiple regression](#sec-multipleRegression), [multicollinearity](#sec-multipleRegressionMulticollinearity) is also a potential concern for mixed models.
Here are the variance inflation factors (VIFs) from an earlier model (VIF is described in @sec-multipleRegressionMulticollinearity):

```{r}
car::vif(pointsPerSeason_positionAgeRandomLinearFixedQuadraticSlopesInteractionExperience)
```
